Wood density and the growth-dependent & independent mortality

### 2015-08-25 ms.Rmd
Updated priors

\begin{equation} \label{eq:alpha0}
\mu_{\log \alpha_{0,s}} \sim \text{normal}\left(-0.87, 0.75\right).\\
\end{equation}

This distribution was derived from data presented by Kobe et al 1995, who estimated the probability of dying at zero growth/2.5 years for juveniles of 10 hardwood species across north-eastern America. Note that these data are also used to define the growth-mortality relationship in SORTIE-ND a spatially explicit forest dynamic model (REF). We converted these probabilities to instantaneous hazard rates by using $\lambda = -\log(1 - Pr(Death))/2.5$, and then log transforming these values and calculating the mean (i.e. -0.87) associated standard error (i.e. 0.75). Based on this prior, we assumed that the average probability of death for an average species was 34%/yr with 95% of species effects occuring between 9% and 84%.

EXAMPLE:
```{r}
# mean species instantaneous hazard rate
M <- c(0.99,0.998,0.555,0.014,0.077,0.999,0.268,0.998,0.985)
k = - log(1-M)/2.5
mean(log(k))

# standard error
sd(log(k))/sqrt(9)
```

@@

  raw_log_c0 ~ normal(0, 1);
  mu_log_c0 ~ normal(-3.86, 0.2);
  sigma_log_c0 ~ cauchy(0, 2.5);

@@

  a2 ~ normal(0,5);
  b2 ~ normal(0,5);
  c2 ~ normal(0,5);
}
```

>>>>

\begin{equation} \label{eq:alpha0}
\mu_{\log \alpha_{0,s}} \sim \text{normal}\left(-0.61, 0.57\right).\\
\end{equation}

This distribution was derived from data presented by Kobe et al 1995, who estimated the probability of dying at zero growth/2.5 years for juveniles of 10 hardwood species across north-eastern America. Note that these data are also used to define the growth-mortality relationship in SORTIE-ND a spatially explicit forest dynamic model (REF). We converted these probabilities to instantaneous hazard rates by using $\lambda = -\log(1 - Pr(Death))/2.5$, and then log transforming these values and calculating the mean (i.e. -0.61) associated standard error (i.e. 0.). Based on this prior, we assumed that the average probability of death for an average species was 42%/yr with 95% of species effects occuring between 16% and 81%.

EXAMPLE:
```{r}
# white ash
median(c(0.999,0.55,0.998))
# sugar maple
median(c(0.998, 0.148, 0.999)
# American beech
median(c(0.014,0.12))
# Yellow Birch
median(c(0.555,0.995))
# Red Maple
0.99
# Eastern hemlock
0.077
# Black cherry
0.998
# White oak
0.82
# Red oak
0.985
#White pine
0.268

# mean species instantaneous hazard rate (used medians for species with multiple measurements)
M <- c(0.998,0.998,0.067,0.775,0.99,0.077,0.998,0.82,0.985, 0.268)
k = - log(1-M)/2.5
mean(log(k))

# standard error
sd(log(k))/sqrt(10)
```

@@

  raw_log_c0 ~ normal(0, 1);
  mu_log_c0 ~ normal(-4.43, 1);
  sigma_log_c0 ~ cauchy(0, 2.5);

@@

  a2 ~ normal(0,5);
  b2 ~ normal(0,5);
  c2 ~ normal(0,5);
}

generated quantities {

  // Declaring fitted parameters
  real a0[n_spp];
  real a1[n_census];

  real b0[n_spp];
  real b1[n_census];

  real c0[n_spp];
  real c1[n_census];

  real alpha_fit;
  real beta_fit;
  real gamma_fit;
  real cumulative_hazard_fit;
  real log_lik_fit;
  real sum_log_lik_fit;

  // Declaring heldout parameters
  real alpha_heldout;
  real beta_heldout;
  real gamma_heldout;
  real cumulative_hazard_heldout;
  real log_lik_heldout[n_obs_heldout];
  real sum_log_lik_heldout;

  // Initialization of summed log likelihoods
  sum_log_lik_fit <- 0;
  sum_log_lik_heldout <- 0;

  // recalulate species random effects
  for (s in 1:n_spp) {
    a0[s] <- exp(raw_log_a0[s] * sigma_log_a0 + mu_log_a0);
    b0[s] <- exp(raw_log_b0[s] * sigma_log_b0 + mu_log_b0);
    c0[s] <- exp(raw_log_c0[s] * sigma_log_c0 + mu_log_c0);
  }

  // recalulate species random effects
  for (t in 1:n_census) {
    a1[t] <- exp(raw_log_a1[t] * sigma_log_a1);
    b1[t] <- exp(raw_log_b1[t] * sigma_log_b1);
    c1[t] <- exp(raw_log_c1[t] * sigma_log_c1);
  }

  // log likelihood for fitted model
  for (i in 1:n_obs) {
    alpha_fit <- a0[spp[i]] * a1[census[i]] * pow(rho_c[spp[i]], a2);
    beta_fit <- b0[spp[i]] * b1[census[i]] * pow(rho_c[spp[i]], b2);
    gamma_fit <- c0[spp[i]] * c1[census[i]] * pow(rho_c[spp[i]], c2);

    cumulative_hazard_fit <- -census_length[i] * (alpha_fit * exp(-beta_fit * growth_dt[i]) + gamma_fit);

    if (y[i] == 0) {
      log_lik_fit <- cumulative_hazard_fit;
    }
    else {
      log_lik_fit <- log1m_exp(cumulative_hazard_fit);
    }
    sum_log_lik_fit <- sum_log_lik_fit + log_lik_fit;
  }

  // log likelihood for held out data
  for (j in 1:n_obs_heldout) {
    alpha_heldout <- a0[spp_heldout[j]] * a1[census_heldout[j]] * pow(rho_c_heldout[spp_heldout[j]], a2);
    beta_heldout <- b0[spp_heldout[j]] * b1[census_heldout[j]] * pow(rho_c_heldout[spp_heldout[j]], b2);
    gamma_heldout <- c0[spp_heldout[j]] * c1[census_heldout[j]] * pow(rho_c_heldout[spp_heldout[j]], c2);

    cumulative_hazard_heldout <- -census_length_heldout[j] * (alpha_heldout * exp(-beta_heldout * growth_dt_heldout[j]) + gamma_heldout);

    if (y_heldout[j] == 0) {
      log_lik_heldout[j] <- cumulative_hazard_heldout;
    }
    else {
      log_lik_heldout[j] <- log1m_exp(cumulative_hazard_heldout);
    }
    sum_log_lik_heldout <- sum_log_lik_heldout + log_lik_heldout[j];
  }
}
```
